---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(npi)
library(tidyverse)

```

# npi
> Access the U.S. National Provider Identifier Registry API
  
  [![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
  [![Travis build status](https://travis-ci.org/frankfarach/npi.svg?branch=master)](https://travis-ci.org/frankfarach/npi)
  [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/frankfarach/npi?branch=master&svg=true)](https://ci.appveyor.com/project/frankfarach/npi)
  [![Coverage status](https://codecov.io/gh/frankfarach/npi/branch/master/graph/badge.svg)](https://codecov.io/github/frankfarach/npi?branch=master)

Provide access to the API for the U.S. National Provider Identifier (NPI) Registry Public Search provided by the Center for Medicare and Medicaid Services (CMS): https://npiregistry.cms.hhs.gov/.

## Installation

This package can be installed directly from this Github repo:

```{r install, eval = FALSE}
devtools::install_github("frankfarach/npi")
library(npi)
```

## Usage

### Searching

`search_npi()` searches the public NPI Registry using the [parameters](https://npiregistry.cms.hhs.gov/registry/help-api) made available by the API and returns the results as a [tidy data frame](http://tibble.tidyverse.org/), or tibble.

```{r search_npi}
# Search for orthopedic individual providers in Atlanta, Georgia
res <- search_npi(provider_type = 1,
                  taxonomy = "Orthoped*",
                  city = "Atlanta",
                  state = "GA")
res
```

### Working with Search Results 

The data returned from `search_npi()` is organized according to its relationship to the NPI (`number` column). Data elements with a 1-to-1 relationship with NPI appear in vector columns, whereas elements with a many-to-1 relationship with NPI exist within [list columns](http://r4ds.had.co.nz/many-models.html#list-columns-1). Each element of a list column is a list of tibbles.

There are three such columns:

 * `taxonomies`: Service classification and licnese information
 * `addresses`: Location and mailing address information
 * `identifiers`: Miscellaneous provider identifiers and credential information

Any of these columns can be extracted as a tidy data frame using `dplyr::unnest()`:

```{r unnest}
tax <- res %>%
  select(number, taxonomies) %>% 
  unnest()
tax
```

By repeating this process with other list columns, you can create multiple tidy data frames and join them to make your own master tidy data frame.

```{r multiple_listcols}
# Create tidy data frame of identifiers
ids <- res %>%
  select(number, identifiers) %>% 
  unnest()
ids

# Join taxonomies and identifiers by `number`
left_join(tax, ids, by = "number")
```

### Paging

By default, `search_npi()` will return at most 200 records per request, which is the maximum set by the API. To return additional records, specify the number of records to skip using the skip argument.

```{r skip}
# Return records 200-400
search_npi(
  city = "San Francisco",
  state = "CA",
  skip = 200
  )
```

### Validating NPIs

Use `is_valid_npi()` to check whether a candidate number is a valid NPI number:

```{r valid_npi_ex}
# Validate one-off NPIs
is_valid_npi(1234567893)
is_valid_npi(1234567898)

```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

MIT (c) [Frank Farach](https://github.com/frankfarach)
