---
title: "Advanced Uses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{advanced-use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(npi)
library(purrr)
```

This vignette provides an brief exploration to advanced uses of the npi package.

## Run `npi_search()` on multiple search terms

Users can serially execute searches for multiple values of a single query parameter by using `npi` in combination with the [`purrr`](https://purrr.tidyverse.org/) package. In the example below, we search multiple NPI numbers. A single tibble is returned with record information corresponding toe matching records.

```{r}
npis <- c(1992708929, 1831192848, 1699778688, 1111111111)  # Last element doesn't exist

out <- npis %>% 
  purrr::map(., ~ npi_search(number = .)) %>% 
  dplyr::bind_rows()

npi_summarize(out)
```

Here we search for multiple zip codes in Los Angeles County.

```{r}
codes <- c(90210, 90211, 90212)

zip_3 <- codes %>% 
  purrr::map(., ~ npi_search(postal_code  = .)) %>% 
  dplyr::bind_rows() 

npi_flatten(zip_3)
```

Consult the R for Data Science [chapter on iteration](https://r4ds.had.co.nz/iteration.html) to learn more about using the `purrr` package.

## Note on NPI Downloadable Files

CMS regularly releases full NPI data files [here](https://download.cms.gov/nppes/NPI_Files.html). Consider downloading the entire data set as needed. Data dissemination files are zipped and will exceed 4GB upon decompression. Additional documentation is provided [here](https://www.cms.gov/Regulations-and-Guidance/Administrative-Simplification/NationalProvIdentStand/DataDissemination).




