---
title: "Advanced Uses"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{advanced-use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(npi)
library(purrr)
```

This vignette explores advanced uses of the npi package.

`npi` is an R package that allows R users to access the [U.S. National Provider Identifier (NPI) Registry](https://npiregistry.cms.hhs.gov/) API by the Center for Medicare and Medicaid Services (CMS). The package makes it easy to obtain administrative data linked to a specific individual or organizational healthcare provider. Additionally, users can perform advanced searches based on provider name, location, type of service, credentials, and many other attributes.

See the npi::npi vignette for an introduction to the package.

## Note on NPI Downloadable Files

CMS regularly releases full NPI data files [here](https://download.cms.gov/nppes/NPI_Files.html). We recommend that users download the data file if they need to work with the entire dataset. The API and `npi_search()` returns a maximum of 1,200 records. Also consider downloading the entire data if you need to work with more than the maximum. Data dissemination files are zipped and will exceed 4GB upon decompression. Additional documentation is provided [here](https://www.cms.gov/Regulations-and-Guidance/Administrative-Simplification/NationalProvIdentStand/DataDissemination).

## Run `npi_search()` on multiple search terms

`npi_search()` enables search for a defined set query parameters. The function is not designed for search on multiple values of the same argument at once, as for example in the case of multiple NPI numbers in a single function call. However, users can still serially execute searches for multiple values of a single query parameter by using `npi` in combination with the [`purrr`](https://purrr.tidyverse.org/) package. In the example below, we search multiple NPI numbers. A single tibble is returned with record information corresponding to matching records. The [purrr:map()](https://purrr.tidyverse.org/reference/map.html) function is used to apply the `npi_search()` function on each element of the vector. Thereafter, the [dplyr::bind_rows()](https://dplyr.tidyverse.org/reference/bind.html) function is used to combine the list of dataframes together into a single dataframe.

```{r}
npis <- c(1992708929, 1831192848, 1699778688, 1111111111)  # Last element doesn't exist

out <- npis %>% 
  purrr::map(., ~ npi_search(number = .)) %>% 
  dplyr::bind_rows()

npi_summarize(out)
```

Here we search for multiple zip codes in Los Angeles County.

```{r}
codes <- c(90210, 90211, 90212)

zip_3 <- codes %>% 
  purrr::map(., ~ npi_search(postal_code  = .)) %>% 
  dplyr::bind_rows() 

npi_flatten(zip_3)
```

Consult the R for Data Science [chapter on iteration](https://r4ds.had.co.nz/iteration.html) to learn more about using the `purrr` package.

Alternatively, you can use a simple for loop instead if you are unfamiliar with the tidyverse approach.

```{r}
npis <- c(1992708929, 1831192848, 1699778688, 1111111111)  # Last element doesn't exist
combined_df  <- data.frame()
for(i in npis){
  combined_df <- rbind(combined_df, npi_search(number = i))
}

npi_summarize(combined_df)
```





